---
title: "Bayesian-thresholding multiple systems estimation"
author: "Bernard W. Silverman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian-thresholding multiple systems estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(MultipleSystemsEstimation)
have_mcmcpack <- requireNamespace("MCMCpack", quietly = TRUE)
```

# Introduction

This vignette describes the Bayesian-threshold approach to multiple systems
estimation developed in Silverman (2020), together with its implementation in
`estimate_population_bayesthresh()`.

Only the aspects specific to Bayesian thresholding are considered here.
General background on multiple systems estimation and Poisson log-linear
models is given elsewhere in the `MultipleSystemsEstimation` package.

The basic Bayesian-threshold procedure described here follows Silverman
(2020). The extension to three-list interactions, and the appropriate use of the
Fienberg-Rinaldo existence check within this Bayesian-threshold
implementation, were not part of the original paper and are additions in
the present implementation.

# The Bayesian-threshold procedure

## Bayesian thresholding

In multiple systems estimation, the choice of interaction terms in the
log-linear model can have a substantial effect on the estimate of the
unobserved population. Bayesian thresholding provides a way of selecting
these interactions using a posterior distribution rather than repeated
significance testing or some other model selection approach.

Initially, consider models which only have main effects and two-list
interactions. The approach begins with a model containing all possible
interactions, fits that model using MCMC, and then removes interactions for
which the posterior evidence of a non-zero effect is weak. The reduced model
is then fitted again to obtain the final population estimate and selected
posterior quantiles.

The intercept and main effects are given independent improper flat
priors. The interaction parameters are independent of these and of one another.
By default, each interaction parameter has a Gaussian prior centred at zero
with variance `prior_variance`; setting `prior = "improper"` instead gives the
interaction parameters improper flat priors on \((-\infty,\infty)\).

For each interaction parameter, the strength of evidence that it differs from
zero is measured by

\[
\frac{|\operatorname{E}(\beta_{ij}\mid\text{data})|}
     {\operatorname{sd}(\beta_{ij}\mid\text{data})}.
\]

Interactions for which this quantity is below the specified threshold are removed, and the model is then fitted again using the retained interactions.

The default threshold is 2. Thus the procedure retains interactions for which the absolute posterior mean is at least two posterior standard deviations from zero.

The method can also be extended to allow three-list interactions while
preserving the hierarchical structure of the log-linear model. The two-list
interactions are thresholded first, as described above. A three-list
interaction is then eligible for consideration only if all three of its
constituent two-list interactions have been retained. The model containing the
retained two-list interactions together with all eligible three-list
interactions is fitted, and the three-list interactions are thresholded using
the same posterior mean to posterior standard deviation ratio. The resulting
hierarchical model is then fitted once more to obtain the final population
estimate and selected posterior quantiles.

For example, a three-list interaction \(ijk\) is eligible
for inclusion only if the interactions \(ij\), \(ik\), and \(jk\) have all
been retained at the first stage. At the second stage, only the eligible
three-list terms are thresholded; the two-list terms that survived the initial
thresholding are kept regardless of their new posterior means. This ensures
that the selected model remains hierarchical. 

## MCMC estimation

Posterior sampling is carried out using `MCMCpack::MCMCpoisson()`, which applies Markov chain Monte Carlo to the Poisson log-linear model. The MCMC run produces posterior draws for the intercept, the main effects, and the interaction parameters in the model. These draws are used both for the thresholding step described above and, after the selected model has been refitted, for inference about population size.

In the log-linear formulation, the intercept determines the fitted mean of the
unobserved cell, corresponding to individuals appearing on none of the lists.
For each posterior draw of the intercept \(\mu\), the corresponding size of
the unobserved cell is
\(
\exp(\mu).
\)

Adding this quantity to the observed population gives a posterior draw for the total population size. The resulting collection of draws therefore gives the posterior distribution of the total population. By default, `estimate_population_bayesthresh()` reports the posterior median as the population estimate, together with selected posterior quantiles. The full set of posterior draws can also be returned if required.

The thresholding step approximates a Bayesian model with a mixed prior for the parameters of an atom of probability at zero and some other symmetric density, hence modelling a parsimonious interaction structure; see
Silverman (2020) for further discussion.

# A basic analysis

The Kosovo data provide a useful illustration because allowing three-list
interactions leads to a genuinely different selected model.

We first fit the Bayesian-threshold procedure using only two-list
interactions.

```{r kosovo-pairwise, eval=have_mcmcpack}
estimate_population_bayesthresh(
    Kosovo,
    maxorder = 2
)
```

With the default proper prior for the interaction parameters, the retained two-list interactions are

EXH:ABA, EXH:OSCE, ABA:OSCE, EXH:HRW, and OSCE:HRW.

The posterior median estimate of the total population is about 14,000. The
2.5% and 97.5% posterior quantiles are about 12,200 and 16,400 respectively.

We can now allow three-list interactions by setting `maxorder = 3`.
For this example we use a threshold of 2.5, rather than the default value of
2, in order to illustrate the thresholding of eligible three-list
interactions.

```{r kosovo-threefactor, eval=have_mcmcpack}

estimate_population_bayesthresh(
    Kosovo,
    maxorder = 3,
    threshold = 2.5
)
```

The same five two-list interactions are retained. Two three-list interactions
are eligible for consideration:

EXH:ABA:OSCE and EXH:OSCE:HRW.

Their threshold statistics are about 4.26 and 2.45 respectively. At a
threshold of 2.5, `EXH:ABA:OSCE` is retained while `EXH:OSCE:HRW` is
thresholded out.  The threshold was only used to illustrate how some of the three-list interactions can be retained and others thresholded.  The default threshold of 2 retains both eligible three-list interactions and yields an estimate of 11,800; the 2.5% and 97.5% posterior quantiles are about 10,000 and 14,600, still substantially different from the two-list result.

### Improper priors

The preceding analysis used proper normal priors for the interaction
parameters. The same thresholding procedure can instead be used with
improper flat priors for the interaction parameters. 

Some additional care is needed because
an interaction whose sufficient statistic is zero has its posterior
distribution concentrated at minus infinity. Such terms are accounted for
before the MCMC fit is carried out.
Effects whose posterior distribution is concentrated at minus infinity
are identified automatically and reported in the
`minus_infinite_estimated_effects` component of the returned object. The
effects are still retained in the model, but are not included in
`retained_interactions`, which reports only retained interaction effects
estimated by MCMC. For example, with the five-list UK data:

```{r uk-improper-minus-infinity, eval=have_mcmcpack}

estimate_population_bayesthresh(
  UKdat_5,
  prior = "improper"
)
```

Here `LA:GP` is retained in the model with its effect estimated as minus
infinity, and is therefore reported in `minus_infinite_estimated_effects`
rather than in `retained_interactions`.

### Checking existence with improper interaction priors

When improper flat priors are used for the interaction parameters, some
additional existence checks are required. For some datasets there are
log-linear models for which there is no maximum likelihood solution even
after allowing some parameters to take the value \(-\infty\). This can be
checked using the criterion of Fienberg and Rinaldo (2012).

With improper priors on the interaction parameters, failure of this
criterion also implies that the posterior distribution is not proper.
The package therefore applies the Fienberg-Rinaldo check when
`prior = "improper"`.

- **Initial two-list fit.** If the model containing all two-list
  interactions fails the Fienberg-Rinaldo check, the procedure stops before
  thresholding the two-list interactions.

- **Three-list extension.** With `maxorder = 3`, the model containing the
  retained two-list interactions and all eligible three-list interactions
  is also checked. If it fails the criterion, the proposed three-list
  extension is not fitted and the completed two-list analysis is returned.

These checks are not required when proper priors are used for the
interaction parameters. Results on posterior propriety for Poisson
log-linear models, including those of Forster (2010), show that proper
priors can yield a proper posterior even when the corresponding maximum
likelihood estimate does not exist. The proper-prior Bayesian procedure
therefore proceeds without applying the Fienberg-Rinaldo criterion.

### Examples with improper priors

The two possible failures can be illustrated with the `Korea` and `Western`
datasets.

For `Korea`, the model containing all two-list interactions fails the
Fienberg-Rinaldo criterion. Thus the improper-prior analysis stops before
the two-list interactions can be thresholded.

```{r korea-existence, error=TRUE, eval=have_mcmcpack}
estimate_population_bayesthresh(
  Korea,
  prior = "improper"
)
```

With the default proper prior, by contrast, the initial Bayesian fit can be
carried out. In this example the interaction `b:d` is subsequently
thresholded out, so that the retained model is no longer one of the models
that fails the Fienberg-Rinaldo criterion.

The second possibility is illustrated by Western. To make the issue
visible, we use the unusually low value `threshold = 0.2`.

```{r western-existence, eval=have_mcmcpack}
estimate_population_bayesthresh(
  Western,
  prior = "improper",
  maxorder = 3,
  threshold = 0.2
)
```
The initial two-list fit and thresholding are completed, but the model
containing the retained two-list interactions and all eligible three-list
interactions fails the Fienberg-Rinaldo criterion. The three-list extension
is therefore not fitted and the completed two-list analysis is returned.

With a proper interaction prior the same Fienberg-Rinaldo check is not
applied, and the three-list fit and thresholding proceed.

# MCMC settings and posterior output

The principal MCMC settings can be controlled through arguments passed to
`MCMCpack::MCMCpoisson()`, including `burnin`, `mcmc`, `thin`, `tune`, and
`seed`. The defaults are intended to be suitable for routine use, although
longer runs may be appropriate for a substantive final analysis.
`MCMCpack::MCMCpoisson()` uses a reproducible default seed when `seed` is not
specified. A different seed can be supplied through `...` if a different MCMC
run is required.

By default, the full posterior sample of total population size is not returned.
It can be retained, when required, by setting `return_posterior = TRUE`.

# Summary

Bayesian thresholding provides a parsimonious way to select interaction effects
using posterior evidence. The implementation allows either proper or improper
priors for interaction effects, can extend the thresholding procedure
hierarchically to three-list interactions, and checks the relevant models
against the Fienberg-Rinaldo existence criterion when improper priors are used
for the interactions. The returned object reports
the selected interaction structure and posterior inference for total population
size, with the full posterior sample available optionally.

# References

Silverman, B. W. (2020).
*Multiple-systems analysis for the quantification of modern slavery:
classical and Bayesian approaches*.
*Journal of the Royal Statistical Society: Series A (Statistics in Society)*, **183**(3), 691--736. doi:10.1111/rssa.12505.

Fienberg, S. E. and Rinaldo, A. (2012).
*Maximum likelihood estimation in log-linear models*.
*The Annals of Statistics*, **40**, 996--1023.

Forster, J. J. (2010). Bayesian inference for Poisson and multinomial
log-linear models. *Statistical Methodology*, **7**, 210--224.

Martin, A. D., Quinn, K. M. and Park, J. H. (2011).
*MCMCpack: Markov Chain Monte Carlo in R*.
*Journal of Statistical Software*, **42**(9), 1--21.
