---
title: "Working with hierarchical models and capture histories"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with hierarchical models and capture histories}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MultipleSystemsEstimation)

# Access to implementation functions used in the technical sections below.
.mse_internal <- function(name) {
  getFromNamespace(name, "MultipleSystemsEstimation")
}
```

# Introduction

Multiple systems estimation uses a binary *capture history* to record the
lists on which an individual appears. Capture histories are also used to
index the parameters in the Poisson log-linear models used to fit the 
observed data and to predict the unobserved *dark figure*.  

The log-linear models used by the
package are hierarchical: whenever an interaction is included, all its
lower-order terms are included as well.

The public interface of `MultipleSystemsEstimation` describes hierarchical
models by readable strings such as

```r
"[12,23]"
```

The package also uses an integer encoding of capture histories and model
terms internally. This vignette first describes the public hierarchy notation
and the functions for enumerating and navigating models. It then explains the
integer encoding and the internal functions built on it. The latter are
documented to make the implementation transparent; they are not part of the
supported public interface. They may also be useful to developers building other methods based on
hierarchical models.

# Hierarchical-model notation

A hierarchical model is specified by its *generators*, enclosed in square
brackets and separated by commas. For example,

```r
"[12,23]"
```

has generators 12 and 23. It therefore includes the interactions between
lists 1 and 2 and between lists 2 and 3, together with the intercept and the
main effects for lists 1, 2 and 3. Lower-order terms implied by the generators
do not need to be written separately. These are the *elements* of the model, so `[12,23]` has elements
\(\varnothing, 1, 2, 3, 12\), and \(23\).

Similarly,

```r
"[123,4]"
```

contains the three-list interaction 123, all its lower-order terms, and the
main effect for list 4.  

## Enumerating hierarchical models

The function `get_hierarchical_models()` returns the available hierarchical models which include
all main effects, for a
specified number of lists and maximum interaction order. The number of models increases
very rapidly as the number of lists increases, and the routine works by selecting from 
a pre-computed internal data file. All models for up to five lists are included, but for six lists
only those of order at most two.  

For example, the
hierarchical models for three lists with interactions of order at most two
are obtained by

```{r enumerate-models}
get_hierarchical_models(nlists = 3, maxorder = 2)
```

The function returns model strings, so its results can be passed directly to
other functions that accept hierarchical-model specifications. 

## Neighbouring models

Two hierarchical models are *neighbours* if one can be obtained from the other
by adding or removing a single element while retaining hierarchy.

The *parents* of a capture history are obtained
by removing one list, and its *children* by adding one list. Thus `13` is a
parent of `123`, while `1235` is a child of `123`.
An *outer neighbour* of a hierarchical model is obtained by adding an element all of whose immediate
parents are already in the model. 

An *inner neighbour* is obtained by removing a single maximal element of the model—that is, an element having no children within the model. These maximal elements are the generators displayed in the hierarchical-model notation. Only the generator itself is removed; all its lower-order elements remain. For example, removing `123` from `[123]` gives `[12,13,23]`. By default, main effects are protected from removal.

The function `find_neighbour_hierarchies()` can return all neighbours or restrict the
search to one direction:

```{r model-neighbours}
model <- "[12,23]"

find_neighbour_hierarchies(model, type = "all")
find_neighbour_hierarchies(model, type = "outer")
find_neighbour_hierarchies(model, type = "inner")
```

The `nlists` argument may be supplied when the full system contains lists not
mentioned in `model`. The `maxorder` argument restricts the interaction order
of outer neighbours, while `keepmaineffects = TRUE` prevents inner moves that
remove a main effect.

# Internal integer encoding

The remainder of this vignette describes the package's internal representation.
Users fitting models through the public interface do not need to construct or
interpret these integers.

For a set of lists \(S\), the encoded capture history is

\[
k = 1 + \sum_{i \in S} 2^{i-1}.
\]

The offset of one makes the all-zero capture history equal to
1 and allows encoded values to be used directly as row or column indices in R.
The same encoding is used for log-linear model terms: 1 is the intercept,
single-list histories represent main effects, and histories involving two or
more lists represent interactions.

For three lists the complete encoding is:

| Lists present | Binary indicators | Encoded value | Model term |
|:--------------|:------------------|--------------:|:-----------|
| none          | 000               | 1             | intercept  |
| 1             | 100               | 2             | main effect 1 |
| 2             | 010               | 3             | main effect 2 |
| 12            | 110               | 4             | interaction 12 |
| 3             | 001               | 5             | main effect 3 |
| 13            | 101               | 6             | interaction 13 |
| 23            | 011               | 7             | interaction 23 |
| 123           | 111               | 8             | interaction 123 |

The binary indicators are displayed in least-significant-bit-first order: the indicator for list 1 appears on the left.

## Encoding and decoding

Internally, `encode_capture()` maps a logical or binary vector to its integer
encoding, while `decode_capture()` performs the reverse operation:

```{r encode-decode}
encode_capture <- .mse_internal("encode_capture")
decode_capture <- .mse_internal("decode_capture")

encode_capture(c(1, 0, 1))
decode_capture(6, nlists = 3)
```

The function `decode_capture()` genuinely needs `nlists`. For example, encoded value 4
represents lists 1 and 2 in both a two-list system and a five-list system, but
the required decoded vectors have different lengths:

```{r decode-length}
decode_capture(4, nlists = 2)
decode_capture(4, nlists = 5)
```

The trailing absent lists cannot be recovered from the encoded value alone.

## Parents, children, ancestors and descendants

Routines are provided for finding parents and children of encoded captures.

An immediate parent is obtained by removing one list. Encoded history 8
represents 123, whose parents are 12, 13 and 23:

```{r parents}
parent_captures <- .mse_internal("parent_captures")

sort(parent_captures(8))
```

The total number of lists is not needed: parents can involve only lists already
present in the capture history. 
The *ancestors* comprise the history itself and everything obtained by
repeatedly taking parents. Thus all encoded values from 1 to 8 are ancestors
of 123:

```{r ancestors}
ancestors <- .mse_internal("ancestors")

ancestors(8)
```
By contrast, an immediate child is obtained by adding one list, and the
possible additions depend on the total number of lists in the system:

```{r children}
child_captures <- .mse_internal("child_captures")

child_captures(4, nlists = 4)
```

Here encoded history 4 represents 12; in a four-list system its children are
123 and 124. 

The same dependence on the complete system applies to
descendants, defined in the natural way by repeatedly finding children:

```{r descendants}
descendants <- .mse_internal("descendants")

descendants(8, nlists = 4)
```

Consequently, `parent_captures()` and `ancestors()` do not require `nlists`,
whereas `child_captures()` and `descendants()` do.

## Converting between hierarchy strings and encoded terms

The package converts a hierarchy string to the complete vector of encoded
terms before fitting a model. For example, the hierarchy generated by 12 and
23 contains the intercept, all three main effects, and the two specified
interactions:

```{r hierarchy-conversion}
convert_from_hierarchy <- .mse_internal("convert_from_hierarchy")
convert_to_hierarchy <- .mse_internal("convert_to_hierarchy")

encoded_model <- convert_from_hierarchy("[12,23]")
sort(encoded_model)

convert_to_hierarchy(encoded_model)
```

The function `convert_to_hierarchy()` raises an error if the encoded captures in its vector
argument do not define a hierarchical model; it does not silently add missing
terms.
For example, encoded value 4 represents the
interaction 12, so the vector `c(1, 4)` is invalid because the two main effects
are absent:

```{r invalid-hierarchy, error=TRUE}
convert_to_hierarchy(c(1, 4))
```

## Boundary terms

For a hierarchical set, a *boundary term* is a term not currently in the
model whose immediate parents are all present. Adding a boundary term therefore
preserves hierarchy.

For the hierarchy generated by 12 and 23 in a three-list system, the missing
pair 13 is on the boundary:

```{r boundary}
boundary_captures <- .mse_internal("boundary_captures")

boundary_captures(encoded_model, nlists = 3)
```

Adding it gives the outer neighbour `[12,13,23]`. The public
`find_neighbour_hierarchies()` function performs this calculation and returns
the resulting readable model strings.

## The master design matrix

The function `make_master_design()` constructs the inclusion matrix mapping log-linear
parameters to observable capture histories. Its columns are the encoded model
terms and its rows are the nonempty capture histories. An entry is one exactly
when the column term is an ancestor of the row history.

```{r master-design}
make_master_design <- .mse_internal("make_master_design")

master <- make_master_design(3)
dim(master)
master[c("2", "8"), ]
```

For three lists there are seven observable capture histories and eight possible
parameters, including the intercept. The row for encoded history 2 depends
only on the intercept and main effect 1, while the row for encoded history 8
depends on every possible term.

# Public interface and internal machinery

The division between the public hierarchy interface and the internal encoding
is summarised below.

| Role | Functions |
|:-----|:----------|
| Public hierarchy strings | `get_hierarchical_models()`, `find_neighbour_hierarchies()` |
| Internal encoding | `encode_capture()`, `decode_capture()` |
| Internal navigation | `parent_captures()`, `child_captures()`, `ancestors()`, `descendants()`, `boundary_captures()` |
| Internal conversion | `convert_from_hierarchy()`, `convert_to_hierarchy()` |
| Internal model construction | `make_master_design()` |

The public functions deliberately use readable hierarchy strings. The encoded
representation remains documented because it explains the structure of the
package and is useful to developers, but it is not required for ordinary use.

# Reference

Silverman, B. W., Chan, L. and Vincent, K. (2024). Bootstrapping multiple
systems estimates to account for model selection. *Statistics and Computing*,
**34**, 44. <https://doi.org/10.1007/s11222-023-10346-9>.
