---
title: "Reproducing Silverman, Chan and Vincent (2024)"
author: "Bernard Silverman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing Silverman, Chan and Vincent (2024)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)

library(MultipleSystemsEstimation)
```

# Introduction

This vignette gives reproducibility code for the numerical results in

Silverman, B. W., Chan, L. and Vincent, K. (2024).
*Bootstrapping multiple systems estimates to account for model selection*,
*Statistics and Computing* **34**, 44.

The calculations use the current `MultipleSystemsEstimation` interface.

The general rule in this vignette is:

* calculations for finite values of `ntop` are carried out directly;
* the paper's value `nboot = 1000` is used for the substantive calculations;
* exhaustive five-list bootstrap calculations over all candidate models are
  not run during an ordinary vignette build, because they are much more
  expensive;
* where an exhaustive result is needed only as a benchmark, the published
  value is inserted explicitly and identified as such;
* code for carrying out the exhaustive calculation is nevertheless shown.

Table 1 of the paper is not reproduced.  It illustrates the issue of
existence of the extended maximum likelihood estimate rather than providing a
numerical application.  The package function `check_identifiability()` provides
the corresponding diagnostic.

# Helper functions

```{r helpers}
pick_ntop <- function(x, values) {
  out <- x$inference[x$inference$ntop %in% values, , drop = FALSE]
  out[match(values, out$ntop), , drop = FALSE]
}

ci_text <- function(lo, hi, rounding = 1) {
  lo <- round(lo / rounding) * rounding
  hi <- round(hi / rounding) * rounding
  paste0("[", lo, ", ", hi, "]")
}

plot_ntop_set <- function(x, at, col, lwd = 1) {
  z <- x[x$ntop %in% at, , drop = FALSE]
  z <- z[match(at, z$ntop), , drop = FALSE]
  xx <- seq_along(at)

  lines(xx, z[["0.025"]], lty = 1, col = col, lwd = lwd)
  lines(xx, z[["0.975"]], lty = 1, col = col, lwd = lwd)

  if (all(c("0.1", "0.9") %in% names(z))) {
    lines(xx, z[["0.1"]], lty = 2, col = col, lwd = lwd)
    lines(xx, z[["0.9"]], lty = 2, col = col, lwd = lwd)
  }
}

plot_two_ntop_bands <- function(x2, est2, x4, est4, at, main,
                                lwd2 = 1, lwd4 = 1) {
  z2 <- x2[x2$ntop %in% at, , drop = FALSE]
  z2 <- z2[match(at, z2$ntop), , drop = FALSE]

  z4 <- x4[x4$ntop %in% at, , drop = FALSE]
  z4 <- z4[match(at, z4$ntop), , drop = FALSE]

  ylim <- range(
    z2[["0.025"]], z2[["0.975"]], est2,
    z4[["0.025"]], z4[["0.975"]], est4,
    finite = TRUE
  )

  xx <- seq_along(at)

  plot(
    xx, rep(NA_real_, length(xx)),
    type = "n", xaxt = "n",
    xlab = "ntop", ylab = "Population estimate",
    ylim = ylim, main = main
  )
  axis(1, at = xx, labels = at)

  plot_ntop_set(x2, at, "blue", lwd = lwd2)
  plot_ntop_set(x4, at, "red", lwd = lwd4)
  abline(h = est2, lty = 4, col = "blue", lwd = lwd2)
  abline(h = est4, lty = 4, col = "red", lwd = lwd4)

  legend(
    "topright",
    legend = c(
      "maximum order 2", "maximum order 4",
      "95% limits", "80% limits", "point estimate"
    ),
    col = c("blue", "red", "black", "black", "black"),
    lty = c(1, 1, 1, 2, 4),
    bty = "n"
  )
}
```

# Table 2: Kosovo

Kosovo has only 113 admissible candidate models, so the exhaustive calculation
is manageable and is carried out directly.

```{r table2-calc}
data(Kosovo)

kos <- vary_ntop_bca(
  Kosovo,
  maxorder = ncol(Kosovo) - 2,
  ntopmax = Inf,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

kos_last <- max(kos$inference$ntop)

kos_tab <- pick_ntop(kos, c(1, 5, 10, kos_last))

table2 <- data.frame(
  ntop = c("1", "5", "10", paste0(kos_last, " (Inf)")),
  `80% CI` = ci_text(kos_tab[["0.1"]], kos_tab[["0.9"]], 100),
  `95% CI` = ci_text(kos_tab[["0.025"]], kos_tab[["0.975"]], 100),
  check.names = FALSE
)

knitr::kable(table2)
```

The published table, rounded to the nearest 100, is:

```{r table2-published}
table2_published <- data.frame(
  ntop = c("1", "5", "10", "113 (Inf)"),
  `80% CI` = c(
    "[9500, 11300]",
    "[8500, 11500]",
    "[7400, 12200]",
    "[7400, 12200]"
  ),
  `95% CI` = c(
    "[9100, 12000]",
    "[8500, 17000]",
    "[6900, 18000]",
    "[6900, 18000]"
  ),
  check.names = FALSE
)

knitr::kable(table2_published)
```

# Table 3: Korea

There are only six admissible models after the existence check, so the
exhaustive calculation is again run directly.

```{r table3-calc}
data(Korea)

kor <- vary_ntop_bca(
  Korea,
  maxorder = 2,
  ntopmax = Inf,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

kor_last <- max(kor$inference$ntop)

kor_tab <- pick_ntop(kor, c(1, 2, kor_last))

table3 <- data.frame(
  ntop = c("1", "2", paste0(kor_last, " (Inf)")),
  `80% CI` = ci_text(kor_tab[["0.1"]], kor_tab[["0.9"]]),
  `95% CI` = ci_text(kor_tab[["0.025"]], kor_tab[["0.975"]]),
  check.names = FALSE
)

knitr::kable(table3)
```

The published intervals are:

```{r table3-published}
table3_published <- data.frame(
  ntop = c("1", "2", "6 (Inf)"),
  `80% CI` = c("[136, 198]", "[135, 286]", "[135, 288]"),
  `95% CI` = c("[131, 248]", "[130, 348]", "[128, 349]"),
  check.names = FALSE
)

knitr::kable(table3_published)
```

# Figure 1: five-list UK data

The finite-`ntop` calculations are performed in the vignette.  We calculate
through `ntop = 60`, which is also sufficient for the finite part of Figure 3.

```{r fig1-calc}
data(UKdat_5)

uk2 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 2,
  ntopmax = 60,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

uk4 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 4,
  ntopmax = 60,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)
```

The paper also shows the exhaustive result at `ntop = Inf`.  Repeating an
exhaustive bootstrap search over all five-list models is deliberately omitted
from an ordinary vignette build.  The finite part of the figure is therefore
drawn directly from the calculations above.

```{r fig1}
at1 <- c(1, 2, 5, 10, 20, 50)

plot_two_ntop_bands(
  pick_ntop(uk2, at1), uk2$estimate,
  pick_ntop(uk4, at1), uk4$estimate,
  at = at1,
  main = "Five-list UK data"
)
```

The published exhaustive 95% intervals, from Table 6 of the paper, are
`[11456, 43894]` for maximum order 2 and `[11168, 56052]` for maximum order 4.
Those values may be added as an `Inf` benchmark if an exact visual reproduction
of the published figure is desired.

A fully fresh exhaustive calculation can be requested by replacing
`ntopmax = 60` with `ntopmax = Inf`, but it is substantially more expensive.

# Figure 2: five-list New Orleans data

Again the finite calculations are performed directly, through `ntop = 60`.

```{r fig2-calc}
data(NewOrl_5)

no2 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 2,
  ntopmax = 60,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

no4 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 4,
  ntopmax = 60,
  degree = 1,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)
```

The two sets of curves are essentially coincident for the finite values shown.

```{r fig2}
at2 <- c(1, 2, 5, 10, 20, 50)

plot_two_ntop_bands(
  pick_ntop(no2, at2), no2$estimate,
  pick_ntop(no4, at2), no4$estimate,
  at = at2,
  main = "Five-list New Orleans data",
  lwd2 = 3,
  lwd4 = 1
)
```

For `ntop = Inf`, the paper reports upper 97.5% limits of 1544 for maximum
order 2 and 1547 for maximum order 4.  The exhaustive calculation is again
not performed in an ordinary vignette build.

# Table 4: location of the bootstrap BIC minimum

Table 4 asks, for each bootstrap sample, where the globally optimal bootstrap
model lies in the original-data BIC ranking.  Thus it intrinsically requires
an exhaustive bootstrap search.

The calculation is therefore shown but not evaluated in an ordinary vignette
build.

```{r table4-code, eval=FALSE}
bootstrap_optimal_ranks <- function(zdat, maxorder,
                                    nboot = 1000, iseed = 1234) {
  z <- MultipleSystemsEstimation:::assemble_bic(
    zdat, maxorder = maxorder
  )

  z <- MultipleSystemsEstimation:::bootstrapcal(
    z, nboot = nboot, iseed = iseed
  )

  original_order <- order(z$res[, "BIC"])

  vapply(seq_len(nboot), function(b) {
    v <- z$bootbic[, b]
    ok <- which(is.finite(v))
    if (!length(ok)) return(NA_integer_)

    best <- ok[which.min(v[ok])]
    match(best, original_order)
  }, integer(1))
}

rank_counts <- function(r, ntop = c(1, 5, 10, 50, 100)) {
  vapply(
    ntop,
    function(k) sum(r <= k, na.rm = TRUE),
    integer(1)
  )
}
```

The published counts, out of 1000 bootstrap samples, are:

```{r table4}
table4 <- data.frame(
  dataset = c(
    "Kosovo",
    "UK (maxorder 2)",
    "UK (maxorder 4)",
    "New Orleans (maxorder 2)",
    "New Orleans (maxorder 4)"
  ),
  `1` = c(375, 216, 156, 357, 357),
  `5` = c(929, 583, 622, 622, 621),
  `10` = c(997, 739, 732, 698, 697),
  `50` = c(1000, 930, 898, 915, 913),
  `100` = c(1000, 954, 943, 958, 956),
  check.names = FALSE
)

knitr::kable(table4)
```

# Table 5: degree-2 ranking for Kosovo

The ordinary BIC ordering is `degree = 1`; the neighbourhood-based ordering
from Section 5.1 of the paper is `degree = 2`.

Kosovo is small enough for both orderings to be calculated exhaustively.

```{r table5-calc}
kos_d2 <- vary_ntop_bca(
  Kosovo,
  maxorder = ncol(Kosovo) - 2,
  ntopmax = Inf,
  degree = 2,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)

kos1_95 <- pick_ntop(kos, c(1, 5, 10, max(kos$inference$ntop)))
kos2_95 <- pick_ntop(kos_d2, c(1, 5, 10, max(kos_d2$inference$ntop)))

table5 <- data.frame(
  ntop = c("1", "5", "5", "10", "10",
           paste0(max(kos$inference$ntop), " (Inf)")),
  degree = c("1, 2", "1", "2", "1", "2", "1, 2"),
  `95% CI` = c(
    ci_text(kos1_95[1, "0.025"], kos1_95[1, "0.975"], 100),
    ci_text(kos1_95[2, "0.025"], kos1_95[2, "0.975"], 100),
    ci_text(kos2_95[2, "0.025"], kos2_95[2, "0.975"], 100),
    ci_text(kos1_95[3, "0.025"], kos1_95[3, "0.975"], 100),
    ci_text(kos2_95[3, "0.025"], kos2_95[3, "0.975"], 100),
    ci_text(kos1_95[4, "0.025"], kos1_95[4, "0.975"], 100)
  ),
  check.names = FALSE
)

knitr::kable(table5)
```

The published values are:

```{r table5-published}
table5_published <- data.frame(
  ntop = c("1", "5", "5", "10", "10", "113 (Inf)"),
  degree = c("1, 2", "1", "2", "1", "2", "1, 2"),
  `95% CI` = c(
    "[9100, 12000]",
    "[7900, 17400]",
    "[8000, 13500]",
    "[6900, 18000]",
    "[8000, 17700]",
    "[6900, 18000]"
  ),
  check.names = FALSE
)

knitr::kable(table5_published)
```

# Figure 3: ordinary and degree-2 rankings

Figure 3 compares degree 1 and degree 2 for the five-list UK and New Orleans
data, allowing interactions through order 4.

The degree-1 finite calculations have already been obtained for Figures 1 and
2.  Only the degree-2 finite calculations are new.

```{r fig3-calc}
uk4_d2 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 4,
  ntopmax = 60,
  degree = 2,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)

no4_d2 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 4,
  ntopmax = 60,
  degree = 2,
  nboot = 1000,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)
```

The two orderings necessarily give the same result at `ntop = Inf`.  The
published exhaustive 95% intervals are therefore used only as benchmark
endpoints in the description below.

```{r fig3, fig.height=8}
plot_degree_comparison <- function(d1, d2, at, main) {
  z1 <- pick_ntop(d1, at)
  z2 <- pick_ntop(d2, at)

  xx <- seq_along(at)

  ylim <- range(
    z1[["0.025"]], z1[["0.975"]],
    z2[["0.025"]], z2[["0.975"]],
    d1$estimate, finite = TRUE
  )

  plot(
    xx, rep(NA_real_, length(xx)),
    type = "n", xaxt = "n",
    xlab = "ntop", ylab = "Population estimate",
    ylim = ylim, main = main
  )
  axis(1, at = xx, labels = at)

  lines(xx, z1[["0.025"]], lty = 1)
  lines(xx, z1[["0.975"]], lty = 1)

  lines(xx, z2[["0.025"]], lty = 3, col = "red")
  lines(xx, z2[["0.975"]], lty = 3, col = "red")

  abline(h = d1$estimate, lty = 2)

  legend(
    "topright",
    legend = c("degree 1", "degree 2", "point estimate"),
    col = c("black", "red", "black"),
    lty = c(1, 3, 2),
    bty = "n"
  )
}

at3 <- c(1, 2, 5, 10, 20, 50, 60)

oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 1))

plot_degree_comparison(
  uk4, uk4_d2, at3,
  main = "Five-list UK data"
)

plot_degree_comparison(
  no4, no4_d2, at3,
  main = "Five-list New Orleans data"
)

par(oldpar)
```

The paper's main conclusion is reproduced by these finite calculations: the
degree-2 ordering requires a larger `ntop` for the UK data, whereas for New
Orleans the two approaches are very similar.  At `ntop = Inf`, the two
orderings coincide.

The published exhaustive 95% intervals are `[11168, 56052]` for the five-list
UK data and `[397, 1511]` for the five-list New Orleans data.

# Figure 4: rank of the bootstrap-optimal model

Figure 4 is intrinsically based on an exhaustive bootstrap search.  For every
bootstrap replication, the globally optimal model is first found among all
candidate models, and then its rank under the two original-data orderings is
recorded.

This is therefore another calculation which is deliberately not evaluated in
an ordinary vignette build.

The corrected current-package code is:

```{r fig4-code, eval=FALSE}
bootstrap_rank_pair <- function(zdat, maxorder,
                                nboot = 1000, iseed = 1234) {
  z <- MultipleSystemsEstimation:::assemble_bic(
    zdat, maxorder = maxorder
  )

  z <- MultipleSystemsEstimation:::bootstrapcal(
    z, nboot = nboot, iseed = iseed
  )

  r1_order <- order(z$res[, "BIC"])

  r2_order <- MultipleSystemsEstimation:::.degree2_order(
    rownames(z$res),
    nlists = ncol(z$xdata) - 1,
    maxorder = z$maxorder
  )

  r1 <- rep(NA_integer_, nboot)
  r2 <- rep(NA_integer_, nboot)

  for (b in seq_len(nboot)) {
    v <- z$bootbic[, b]
    ok <- which(is.finite(v))
    if (!length(ok)) next

    best <- ok[which.min(v[ok])]
    r1[b] <- match(best, r1_order)
    r2[b] <- match(best, r2_order)
  }

  keep <- is.finite(r1) & is.finite(r2) & r1 > 1

  data.frame(
    ordinary = r1[keep],
    degree2 = r2[keep]
  )
}

kos_rank <- bootstrap_rank_pair(
  Kosovo,
  maxorder = ncol(Kosovo) - 2
)

no2_rank <- bootstrap_rank_pair(NewOrl_5, maxorder = 2)
no4_rank <- bootstrap_rank_pair(NewOrl_5, maxorder = 4)
uk2_rank <- bootstrap_rank_pair(UKdat_5, maxorder = 2)
uk4_rank <- bootstrap_rank_pair(UKdat_5, maxorder = 4)

rank_sets <- list(
  Kos = kos_rank,
  NO2 = no2_rank,
  NO4 = no4_rank,
  UK2 = uk2_rank,
  UK4 = uk4_rank
)

oldpar <- par(no.readonly = TRUE)
par(mfrow = c(1, 5), mar = c(6, 3, 2, 1))

for (nm in names(rank_sets)) {
  x <- rank_sets[[nm]]

  boxplot(
    list(BIC = x$ordinary, degree2 = x$degree2),
    log = "y",
    main = nm,
    las = 2,
    ylab = if (nm == "Kos") "BIC ranks" else ""
  )
}

par(oldpar)
```

The published Figure 4 shows that the ordinary BIC ranking tends to place the
bootstrap-optimal models substantially lower than the degree-2 ordering.  This
supports the paper's conclusion that the degree-2 neighbourhood construction
does not improve the computational approximation.

# Table 6: downhill search

Section 5.2 compares a greedy downhill search with exhaustive model
enumeration.

For datasets for which the downhill calculation is reasonably fast, it can be
run directly.  For example:

```{r table6-korea}
downhill_korea <- downhill_funs(
  Korea,
  maxorder = 2,
  nboot = 1000,
  iseed = 1234
)

downhill_korea
```

The published Table 6 is:

```{r table6}
table6 <- data.frame(
  dataset = c(
    "UK5 all models",
    "UK5 maxorder 2",
    "UK all models",
    "UK maxorder 2",
    "NewOrl5 all models",
    "NewOrl5 maxorder 2",
    "NewOrl all models",
    "NewOrl maxorder 2",
    "Kosovo all lists",
    "Kosovo maxorder 2",
    "Korea"
  ),
  `downhill estimate` = c(
    12262, 12262, 12350, 12350, 981, 981,
    1110, 1110, 10357, 14342, 157
  ),
  `downhill 95% CI` = c(
    "[9330, 32115]",
    "[9332, 28954]",
    "[9622, 27889]",
    "[9643, 25508]",
    "[397, 1564]",
    "[397, 1564]",
    "[586, 1512]",
    "[586, 1512]",
    "[7133, 18681]",
    "[12525, 16477]",
    "[128, 349]"
  ),
  `all-model estimate` = c(
    "25311", "22991", "six lists", "", "981", "981",
    "eight lists", "", "10357", "14342", "157"
  ),
  `all-model 95% CI` = c(
    "[11168, 56052]",
    "[11456, 43894]",
    "",
    "",
    "[397, 1511]",
    "[397, 1505]",
    "",
    "",
    "[6947, 18010]",
    "[12451, 16273]",
    "[128, 349]"
  ),
  check.names = FALSE
)

knitr::kable(table6)
```

The special five-list UK downhill result in the paper used several additional
order-2 starting models because the main-effects start can converge to a
secondary local minimum.  The published values are retained here rather than
embedding that specialised multiple-start reproduction machinery in the main
vignette.

# Summary

The main computational message of the paper is visible directly from the
finite-`ntop` calculations: confidence intervals widen substantially when model
selection is incorporated, but relatively small values of `ntop` generally
give results close to the exhaustive calculation.

The current package therefore makes it possible to reproduce the substantive
finite-`ntop` calculations directly while keeping exhaustive five-list
bootstrap searches out of ordinary vignette builds.
