---
title: "Reproducing Silverman, Chan and Vincent (2024)"
author: "Bernard Silverman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing Silverman, Chan and Vincent (2024)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(MultipleSystemsEstimation)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# Keep the installed vignette quick enough for an ordinary package build.
# Set this to TRUE for the fuller finite-ntop calculations used in the paper.
full_reproduction <- FALSE

nboot_vignette <- if (full_reproduction) 1000L else 20L
ntop_vignette <- if (full_reproduction) 60L else 20L
ntop_degree2_vignette <- if (full_reproduction) 60L else 10L
```

# Introduction

This vignette gives reproducibility code for the numerical results in

Silverman, B. W., Chan, L. and Vincent, K. (2024). *Bootstrapping multiple systems estimates to account for model selection*, *Statistics and Computing* **34**, 44.

The calculations use the current `MultipleSystemsEstimation` interface.

There are two intended ways to use the vignette.  In the version built with the package, `full_reproduction` is `FALSE`.  This uses `nboot = 20` and restricts the more expensive five-list finite searches to modest values of `ntop`.  These settings are enough to exercise the current code, produce the figures and tables in reduced-precision form, and keep an ordinary package or CRAN vignette build reasonably quick.  They are not intended to reproduce the published Monte Carlo values exactly.

For a closer numerical reproduction of the paper, change

```r
full_reproduction <- TRUE
```

in the setup chunk.  This uses `nboot = 1000`, the value used for the paper, and extends the finite five-list calculations to `ntop = 60`.  Calculations requiring an exhaustive search over all five-list candidate models remain deliberately excluded from an ordinary vignette build; where those results are required, the published values are given and the code needed to recompute them is shown with `eval = FALSE`.  Such exhaustive calculations can take substantially longer.

Table 1 of the paper is not reproduced. It illustrates the issue of existence of the extended maximum likelihood estimate rather than providing a numerical application. The package function `check_identifiability()` provides the corresponding diagnostic.

# Helper functions

```{r helpers}
pick_ntop <- function(x, values) {
  out <- x$inference[x$inference$ntop %in% values, , drop = FALSE]
  out[match(values, out$ntop), , drop = FALSE]
}

ci_text <- function(lo, hi, rounding = 1) {
  lo <- round(lo / rounding) * rounding
  hi <- round(hi / rounding) * rounding
  paste0("[", lo, ", ", hi, "]")
}

finite_at <- function(max_ntop) {
  z <- c(1, 2, 5, 10, 20, 50, 60)
  z[z <= max_ntop]
}

plot_ntop_set <- function(x, at, col, lwd = 1) {
  z <- x[x$ntop %in% at, , drop = FALSE]
  z <- z[match(at, z$ntop), , drop = FALSE]
  xx <- seq_along(at)

  lines(xx, z[["0.025"]], lty = 1, col = col, lwd = lwd)
  lines(xx, z[["0.975"]], lty = 1, col = col, lwd = lwd)

  if (all(c("0.1", "0.9") %in% names(z))) {
    lines(xx, z[["0.1"]], lty = 2, col = col, lwd = lwd)
    lines(xx, z[["0.9"]], lty = 2, col = col, lwd = lwd)
  }
}

plot_two_ntop_bands <- function(x2, est2, x4, est4, at, main,
                                lwd2 = 1, lwd4 = 1) {
  z2 <- x2[x2$ntop %in% at, , drop = FALSE]
  z2 <- z2[match(at, z2$ntop), , drop = FALSE]

  z4 <- x4[x4$ntop %in% at, , drop = FALSE]
  z4 <- z4[match(at, z4$ntop), , drop = FALSE]

  ylim <- range(
    z2[["0.025"]], z2[["0.975"]], est2,
    z4[["0.025"]], z4[["0.975"]], est4,
    finite = TRUE
  )

  xx <- seq_along(at)

  plot(
    xx, rep(NA_real_, length(xx)),
    type = "n", xaxt = "n",
    xlab = "ntop", ylab = "Population estimate",
    ylim = ylim, main = main
  )
  axis(1, at = xx, labels = at)

  plot_ntop_set(x2, at, "blue", lwd = lwd2)
  plot_ntop_set(x4, at, "red", lwd = lwd4)
  abline(h = est2, lty = 4, col = "blue", lwd = lwd2)
  abline(h = est4, lty = 4, col = "red", lwd = lwd4)

}

legend_two_ntop_bands <- function() {
  plot.new()
  legend(
    "center",
    legend = c(
      "maximum order 2", "maximum order 4",
      "95% limits", "80% limits", "point estimate"
    ),
    col = c("blue", "red", "black", "black", "black"),
    lty = c(1, 1, 1, 2, 4),
    ncol = 2,
    cex = 0.9,
    bty = "n"
  )
}
```

# Table 2: Kosovo

Kosovo has only 113 admissible candidate models, so the exhaustive calculation over candidate models is manageable and is carried out directly.  With the ordinary vignette settings the bootstrap uses only 20 replications; set `full_reproduction <- TRUE` to use the paper's `nboot = 1000`.

```{r table2-kosovo}
data(Kosovo)

kos <- vary_ntop_bca(
  Kosovo,
  maxorder = ncol(Kosovo) - 2,
  ntopmax = Inf,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

kos_last <- max(kos$inference$ntop)
kos_tab <- pick_ntop(kos, c(1, 5, 10, kos_last))

table2 <- data.frame(
  ntop = c("1", "5", "10", paste0(kos_last, " (Inf)")),
  `80% CI` = ci_text(kos_tab[["0.1"]], kos_tab[["0.9"]], 100),
  `95% CI` = ci_text(kos_tab[["0.025"]], kos_tab[["0.975"]], 100),
  check.names = FALSE
)

knitr::kable(table2)
```

The published table, rounded to the nearest 100, is:

```{r table2-published}
table2_published <- data.frame(
  ntop = c("1", "5", "10", "113 (Inf)"),
  `80% CI` = c(
    "[9500, 11300]",
    "[8500, 11500]",
    "[7400, 12200]",
    "[7400, 12200]"
  ),
  `95% CI` = c(
    "[9100, 12000]",
    "[8500, 17000]",
    "[6900, 18000]",
    "[6900, 18000]"
  ),
  check.names = FALSE
)

knitr::kable(table2_published)
```

# Table 3: Korea

There are only six admissible models after the existence check, so the exhaustive calculation is again run directly.  As for Table 2, `full_reproduction <- TRUE` changes the bootstrap from 20 to 1000 replications.

```{r table3-korea}
data(Korea)

kor <- vary_ntop_bca(
  Korea,
  maxorder = 2,
  ntopmax = Inf,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

kor_last <- max(kor$inference$ntop)
kor_tab <- pick_ntop(kor, c(1, 2, kor_last))

table3 <- data.frame(
  ntop = c("1", "2", paste0(kor_last, " (Inf)")),
  `80% CI` = ci_text(kor_tab[["0.1"]], kor_tab[["0.9"]]),
  `95% CI` = ci_text(kor_tab[["0.025"]], kor_tab[["0.975"]]),
  check.names = FALSE
)

knitr::kable(table3)
```

The published intervals are:

```{r table3-published}
table3_published <- data.frame(
  ntop = c("1", "2", "6 (Inf)"),
  `80% CI` = c("[136, 198]", "[135, 286]", "[135, 288]"),
  `95% CI` = c("[131, 248]", "[130, 348]", "[128, 349]"),
  check.names = FALSE
)

knitr::kable(table3_published)
```

# Figure 1: five-list UK data

The finite-`ntop` calculations are performed directly.  In an ordinary vignette build they are calculated through `ntop = 20`; with `full_reproduction <- TRUE` they are calculated through `ntop = 60`, with `nboot = 1000`, as used for the finite part of the paper's figure.

```{r fig1-calc}
data(UKdat_5)

uk2 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 2,
  ntopmax = ntop_vignette,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

uk4 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 4,
  ntopmax = ntop_vignette,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)
```

The paper also shows the exhaustive result at `ntop = Inf`. Repeating an exhaustive bootstrap search over all five-list models is deliberately omitted from an ordinary vignette build.  The legend is placed in a separate strip below the plotting region so that it does not obscure the confidence-limit curves.

```{r fig1, fig.height=5.5}
at1 <- finite_at(ntop_vignette)

oldpar <- par(no.readonly = TRUE)
layout(matrix(c(1, 2), nrow = 2), heights = c(5, 1.6))

par(mar = c(4.5, 4.5, 3, 1) + 0.1)
plot_two_ntop_bands(
  pick_ntop(uk2, at1), uk2$estimate,
  pick_ntop(uk4, at1), uk4$estimate,
  at = at1,
  main = "Five-list UK data"
)

par(mar = c(0, 0, 0, 0))
legend_two_ntop_bands()

layout(1)
par(oldpar)
```

The published exhaustive 95% intervals, from Table 6 of the paper, are `[11456, 43894]` for maximum order 2 and `[11168, 56052]` for maximum order 4. Those values may be added as an `Inf` benchmark for an exact visual reproduction.

To recompute an exhaustive endpoint from scratch, replace `ntopmax = ntop_vignette` in the relevant call by `ntopmax = Inf` and use `nboot = 1000`.  This can be substantially more expensive than the finite calculation.

# Figure 2: five-list New Orleans data

Again the finite calculations are performed directly, through `ntop = 20` for the ordinary vignette and through `ntop = 60` when `full_reproduction <- TRUE`.

```{r fig2-calc}
data(NewOrl_5)

no2 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 2,
  ntopmax = ntop_vignette,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)

no4 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 4,
  ntopmax = ntop_vignette,
  degree = 1,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.1, 0.9, 0.975)
)
```

The two sets of curves are essentially coincident for the finite values shown.

```{r fig2, fig.height=5.5}
at2 <- finite_at(ntop_vignette)

oldpar <- par(no.readonly = TRUE)
layout(matrix(c(1, 2), nrow = 2), heights = c(5, 1.6))

par(mar = c(4.5, 4.5, 3, 1) + 0.1)
plot_two_ntop_bands(
  pick_ntop(no2, at2), no2$estimate,
  pick_ntop(no4, at2), no4$estimate,
  at = at2,
  main = "Five-list New Orleans data",
  lwd2 = 3,
  lwd4 = 1
)

par(mar = c(0, 0, 0, 0))
legend_two_ntop_bands()

layout(1)
par(oldpar)
```

For `ntop = Inf`, the paper reports upper 97.5% limits of 1544 for maximum order 2 and 1547 for maximum order 4. The exhaustive calculation is again not performed in an ordinary vignette build.  To recompute it, use `ntopmax = Inf` and `nboot = 1000` in the relevant call.

# Table 4: location of the bootstrap BIC minimum

Table 4 asks, for each bootstrap sample, where the globally optimal bootstrap model lies in the original-data BIC ranking. Thus it intrinsically requires an exhaustive bootstrap search. The calculation is therefore shown but not evaluated in an ordinary vignette build.

The following code carries out the calculation.  It uses 1000 bootstrap samples, as in the paper.

```{r table4-calc, eval=FALSE}
bootstrap_optimal_ranks <- function(zdat, maxorder,
                                    nboot = 1000, iseed = 1234) {
  z <- MultipleSystemsEstimation:::assemble_bic(
    zdat, maxorder = maxorder
  )

  z <- MultipleSystemsEstimation:::bootstrapcal(
    z, nboot = nboot, iseed = iseed
  )

  original_order <- order(z$res[, "BIC"])

  vapply(seq_len(nboot), function(b) {
    v <- z$bootbic[, b]
    ok <- which(is.finite(v))
    if (!length(ok)) return(NA_integer_)

    best <- ok[which.min(v[ok])]
    match(best, original_order)
  }, integer(1))
}

rank_counts <- function(r, ntop = c(1, 5, 10, 50, 100)) {
  vapply(
    ntop,
    function(k) sum(r <= k, na.rm = TRUE),
    integer(1)
  )
}

ranks4 <- list(
  Kosovo = bootstrap_optimal_ranks(
    Kosovo, maxorder = ncol(Kosovo) - 2
  ),
  UK2 = bootstrap_optimal_ranks(UKdat_5, maxorder = 2),
  UK4 = bootstrap_optimal_ranks(UKdat_5, maxorder = 4),
  NewOrleans2 = bootstrap_optimal_ranks(NewOrl_5, maxorder = 2),
  NewOrleans4 = bootstrap_optimal_ranks(NewOrl_5, maxorder = 4)
)

sapply(ranks4, rank_counts)
```

The published counts, out of 1000 bootstrap samples, are:

```{r table4}
table4 <- data.frame(
  dataset = c(
    "Kosovo",
    "UK (maxorder 2)",
    "UK (maxorder 4)",
    "New Orleans (maxorder 2)",
    "New Orleans (maxorder 4)"
  ),
  `1` = c(375, 216, 156, 357, 357),
  `5` = c(929, 583, 622, 622, 621),
  `10` = c(997, 739, 732, 698, 697),
  `50` = c(1000, 930, 898, 915, 913),
  `100` = c(1000, 954, 943, 958, 956),
  check.names = FALSE
)

knitr::kable(table4)
```

# Table 5: degree-2 ranking for Kosovo

The ordinary BIC ordering is `degree = 1`; the neighbourhood-based ordering from Section 5.1 of the paper is `degree = 2`.

Kosovo is small enough for both orderings to be calculated exhaustively.  The ordinary vignette uses 20 bootstrap replications; `full_reproduction <- TRUE` uses 1000.

```{r table5-kosovo}
kos_d2 <- vary_ntop_bca(
  Kosovo,
  maxorder = ncol(Kosovo) - 2,
  ntopmax = Inf,
  degree = 2,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)

kos1_95 <- pick_ntop(kos, c(1, 5, 10, max(kos$inference$ntop)))
kos2_95 <- pick_ntop(kos_d2, c(1, 5, 10, max(kos_d2$inference$ntop)))

table5 <- data.frame(
  ntop = c("1", "5", "5", "10", "10",
           paste0(max(kos$inference$ntop), " (Inf)")),
  degree = c("1, 2", "1", "2", "1", "2", "1, 2"),
  `95% CI` = c(
    ci_text(kos1_95[1, "0.025"], kos1_95[1, "0.975"], 100),
    ci_text(kos1_95[2, "0.025"], kos1_95[2, "0.975"], 100),
    ci_text(kos2_95[2, "0.025"], kos2_95[2, "0.975"], 100),
    ci_text(kos1_95[3, "0.025"], kos1_95[3, "0.975"], 100),
    ci_text(kos2_95[3, "0.025"], kos2_95[3, "0.975"], 100),
    ci_text(kos1_95[4, "0.025"], kos1_95[4, "0.975"], 100)
  ),
  check.names = FALSE
)

knitr::kable(table5)
```

The published values are:

```{r table5-published}
table5_published <- data.frame(
  ntop = c("1", "5", "5", "10", "10", "113 (Inf)"),
  degree = c("1, 2", "1", "2", "1", "2", "1, 2"),
  `95% CI` = c(
    "[9100, 12000]",
    "[7900, 17400]",
    "[8000, 13500]",
    "[6900, 18000]",
    "[8000, 17700]",
    "[6900, 18000]"
  ),
  check.names = FALSE
)

knitr::kable(table5_published)
```

# Figure 3: ordinary and degree-2 rankings

Figure 3 compares degree 1 and degree 2 for the five-list UK and New Orleans data, allowing interactions through order 4.  The degree-1 calculations have already been obtained for Figures 1 and 2. Only the degree-2 calculations are new.

The degree-2 ordering is noticeably more expensive.  The ordinary vignette therefore calculates it only through `ntop = 10`, with 20 bootstrap replications.  Setting `full_reproduction <- TRUE` changes this to `ntop = 60` and `nboot = 1000` for a closer reproduction of the finite part of Figure 3.

```{r fig3-calc}
uk4_d2 <- vary_ntop_bca(
  UKdat_5,
  maxorder = 4,
  ntopmax = ntop_degree2_vignette,
  degree = 2,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)

no4_d2 <- vary_ntop_bca(
  NewOrl_5,
  maxorder = 4,
  ntopmax = ntop_degree2_vignette,
  degree = 2,
  nboot = nboot_vignette,
  iseed = 1234,
  alpha = c(0.025, 0.975)
)
```

The two orderings necessarily give the same result at `ntop = Inf`. The published exhaustive 95% intervals are therefore used only as benchmark endpoints in the description below.

```{r fig3, fig.height=8}
plot_degree_comparison <- function(d1, d2, at, main) {
  z1 <- pick_ntop(d1, at)
  z2 <- pick_ntop(d2, at)

  xx <- seq_along(at)

  ylim <- range(
    z1[["0.025"]], z1[["0.975"]],
    z2[["0.025"]], z2[["0.975"]],
    d1$estimate, finite = TRUE
  )

  plot(
    xx, rep(NA_real_, length(xx)),
    type = "n", xaxt = "n",
    xlab = "ntop", ylab = "Population estimate",
    ylim = ylim, main = main
  )
  axis(1, at = xx, labels = at)

  lines(xx, z1[["0.025"]], lty = 1)
  lines(xx, z1[["0.975"]], lty = 1)

  lines(xx, z2[["0.025"]], lty = 3, col = "red")
  lines(xx, z2[["0.975"]], lty = 3, col = "red")

  abline(h = d1$estimate, lty = 2)
}

legend_degree_comparison <- function() {
  plot.new()
  legend(
    "center",
    legend = c("degree 1", "degree 2", "point estimate"),
    col = c("black", "red", "black"),
    lty = c(1, 3, 2),
    ncol = 1,
    cex = 0.9,
    bty = "n"
  )
}

at3 <- finite_at(ntop_degree2_vignette)

oldpar <- par(no.readonly = TRUE)
layout(
  matrix(c(1, 2, 3), nrow = 3),
  heights = c(1, 1, 0.45)
)

par(mar = c(4.5, 4.5, 3, 1) + 0.1)
plot_degree_comparison(
  uk4, uk4_d2, at3,
  main = "Five-list UK data"
)

par(mar = c(4.5, 4.5, 3, 1) + 0.1)
plot_degree_comparison(
  no4, no4_d2, at3,
  main = "Five-list New Orleans data"
)

par(mar = c(0, 0, 0, 0))
legend_degree_comparison()

layout(1)
par(oldpar)
```

The paper's main conclusion is reproduced by these finite calculations: the degree-2 ordering requires a larger `ntop` for the UK data, whereas for New Orleans the two approaches are very similar. At `ntop = Inf`, the two orderings coincide.

The published exhaustive 95% intervals are `[11168, 56052]` for the five-list UK data and `[397, 1511]` for the five-list New Orleans data.

# Figure 4: rank of the bootstrap-optimal model

Figure 4 is intrinsically based on an exhaustive bootstrap search. For every bootstrap replication, the globally optimal model is first found among all candidate models, and then its rank under the two original-data orderings is recorded.

This is therefore another calculation which is deliberately not evaluated in an ordinary vignette build.  The code below uses 1000 bootstrap samples, as in the paper; it can be run explicitly by a reader who wants the full reproduction.

```{r fig4-calc, eval=FALSE}
bootstrap_rank_pair <- function(zdat, maxorder,
                                nboot = 1000, iseed = 1234) {
  z <- MultipleSystemsEstimation:::assemble_bic(
    zdat, maxorder = maxorder
  )

  z <- MultipleSystemsEstimation:::bootstrapcal(
    z, nboot = nboot, iseed = iseed
  )

  r1_order <- order(z$res[, "BIC"])

  r2_order <- MultipleSystemsEstimation:::.degree2_order(
    rownames(z$res),
    nlists = ncol(z$xdata) - 1,
    maxorder = z$maxorder
  )

  r1 <- rep(NA_integer_, nboot)
  r2 <- rep(NA_integer_, nboot)

  for (b in seq_len(nboot)) {
    v <- z$bootbic[, b]
    ok <- which(is.finite(v))
    if (!length(ok)) next

    best <- ok[which.min(v[ok])]
    r1[b] <- match(best, r1_order)
    r2[b] <- match(best, r2_order)
  }

  keep <- is.finite(r1) & is.finite(r2) & r1 > 1

  data.frame(
    ordinary = r1[keep],
    degree2 = r2[keep]
  )
}

kos_rank <- bootstrap_rank_pair(
  Kosovo,
  maxorder = ncol(Kosovo) - 2
)

no2_rank <- bootstrap_rank_pair(NewOrl_5, maxorder = 2)
no4_rank <- bootstrap_rank_pair(NewOrl_5, maxorder = 4)
uk2_rank <- bootstrap_rank_pair(UKdat_5, maxorder = 2)
uk4_rank <- bootstrap_rank_pair(UKdat_5, maxorder = 4)

rank_sets <- list(
  Kos = kos_rank,
  NO2 = no2_rank,
  NO4 = no4_rank,
  UK2 = uk2_rank,
  UK4 = uk4_rank
)

oldpar <- par(no.readonly = TRUE)
par(mfrow = c(1, 5), mar = c(6, 3, 2, 1))

for (nm in names(rank_sets)) {
  x <- rank_sets[[nm]]

  boxplot(
    list(BIC = x$ordinary, degree2 = x$degree2),
    log = "y",
    main = nm,
    las = 2,
    ylab = if (nm == "Kos") "BIC ranks" else ""
  )
}

par(oldpar)
```

The published Figure 4 shows that the ordinary BIC ranking tends to place the bootstrap-optimal models substantially lower than the degree-2 ordering. This supports the paper's conclusion that the degree-2 neighbourhood construction does not improve the computational approximation.

# Table 6: downhill search

Section 5.2 compares a greedy downhill search with exhaustive model enumeration.

For datasets for which the downhill calculation is reasonably fast, it can be run directly. For example, the Korea calculation can be carried out with the current separate downhill routines as follows.  The ordinary vignette uses 20 bootstrap replications; setting `full_reproduction <- TRUE` uses 1000.

```{r table6-korea}
downhill_korea <- list(
  estimate = downhill_fit(
    counts = Korea[, ncol(Korea)],
    desmat = Korea[, -ncol(Korea), drop = FALSE],
    maxorder = 2
  ),
  bootreps = downhill_bootstrapcal(
    Korea,
    nboot = nboot_vignette,
    iseed = 1234,
    maxorder = 2
  ),
  ahat = downhill_jackknifecal(
    Korea,
    maxorder = 2
  )
)

c(
  estimate = downhill_korea$estimate,
  acceleration = downhill_korea$ahat
)
```

The published Table 6 is:

```{r table6}
table6 <- data.frame(
  dataset = c(
    "UK5 all models",
    "UK5 maxorder 2",
    "UK all models",
    "UK maxorder 2",
    "NewOrl5 all models",
    "NewOrl5 maxorder 2",
    "NewOrl all models",
    "NewOrl maxorder 2",
    "Kosovo all lists",
    "Kosovo maxorder 2",
    "Korea"
  ),
  `downhill estimate` = c(
    12262, 12262, 12350, 12350, 981, 981,
    1110, 1110, 10357, 14342, 157
  ),
  `downhill 95% CI` = c(
    "[9330, 32115]",
    "[9332, 28954]",
    "[9622, 27889]",
    "[9643, 25508]",
    "[397, 1564]",
    "[397, 1564]",
    "[586, 1512]",
    "[586, 1512]",
    "[7133, 18681]",
    "[12525, 16477]",
    "[128, 349]"
  ),
  `all-model estimate` = c(
    "25311", "22991", "six lists", "", "981", "981",
    "eight lists", "", "10357", "14342", "157"
  ),
  `all-model 95% CI` = c(
    "[11168, 56052]",
    "[11456, 43894]",
    "",
    "",
    "[397, 1511]",
    "[397, 1505]",
    "",
    "",
    "[6947, 18010]",
    "[12451, 16273]",
    "[128, 349]"
  ),
  check.names = FALSE
)

knitr::kable(table6)
```

The special five-list UK downhill result in the paper used several additional order-2 starting models because the main-effects start can converge to a secondary local minimum. The published values are retained here rather than embedding that specialised multiple-start reproduction machinery in the main vignette.

For the Korea row, setting `full_reproduction <- TRUE` gives the paper-scale bootstrap calculation with `nboot = 1000`.  Reproducing every row of Table 6 from scratch, particularly the specialised five-list UK calculation, is intentionally outside the computations run during an ordinary vignette build.

# Summary

The main computational message of the paper is visible directly from the finite-`ntop` calculations: confidence intervals widen substantially when model selection is incorporated, but relatively small values of `ntop` generally give results close to the exhaustive calculation.

The default vignette settings are deliberately chosen for package checking rather than Monte Carlo accuracy.  Setting `full_reproduction <- TRUE` restores `nboot = 1000` and the larger finite `ntop` ranges used for the paper.  The few calculations that require exhaustive five-list bootstrap searches are shown separately and are not run automatically.
