---
title: "Reproducing Chan, Silverman and Vincent (2021)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing Chan, Silverman and Vincent (2021)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MultipleSystemsEstimation)
```

# Introduction

Chan, Silverman and Vincent (2021), *Multiple Systems Estimation for Sparse Capture Data: Inferential Challenges When There Are Nonoverlapping Lists*, develops methodology for multiple systems estimation when the observed capture table is sparse.

The paper has two principal themes. First, sparse tables require explicit attention to the existence and identifiability of maximum likelihood estimates, including the possibility of extended maximum likelihood estimates in which some log-linear parameters are equal to \(-\infty\). Second, the paper develops a forward stepwise model-selection procedure and uses the BCa bootstrap to obtain population-size confidence intervals that allow for the model-selection process.

The present vignette shows how the principal numerical results and procedures in the paper can be reproduced using `MultipleSystemsEstimation`.

The package has developed since the original paper was published. Where possible, this vignette uses the current public interface. Some compact internal functions are retained where they are useful for reproducing particular calculations, but the original simulation and supplementary routines are not all carried forward into `MultipleSystemsEstimation`. For full reproduction of the original computational material, including routines specific to the paper, see the now-deprecated `SparseMSE` package. Internal functions used below should not be taken as the recommended interface for new analyses; see the main user guide for current usage.

# Sparse capture data and extended maximum likelihood

Suppose there are \(t\) lists. For a capture history \(\omega\), write \(N_\omega\) for the number of individuals observed on exactly that combination of lists, and

\[
N^*_\omega =
\sum_{\psi\supseteq\omega,\;\psi\ne\emptyset}N_\psi
\]

for the number observed on all lists in \(\omega\), whether or not they also occur on other lists.

A pair of lists \(i,j\) is nonoverlapping if \(N^*_{ij}=0\).

Under the Poisson log-linear model,

\[
N_\omega\sim{\rm Poisson}(\mu_\omega),
\qquad
\log\mu_\omega=\sum_{\theta\subseteq\omega}\alpha_\theta .
\]

If a nonoverlapping pair \(i,j\) is included as an interaction in the model, the extended maximum likelihood estimate of \(\alpha_{ij}\) is \(-\infty\). The package handles this boundary value explicitly rather than relying on numerical iteration towards an increasingly negative value.

The treatment builds on the general theory of Fienberg and Rinaldo (2012), specialised to multiple systems estimation in Chan, Silverman and Vincent (2021).

# Existence and identifiability: Tables 2 and 3

Section 2.6 of the paper distinguishes two problems:

1. the maximum likelihood estimate may fail to exist, even allowing extended parameter values; and
2. the parameters may be nonidentifiable.

These are illustrated by the artificial three-list dataset in Table 2, which is supplied with the package as `Artificial_3`.

```{r}
data(Artificial_3)
Artificial_3
```

The current function `check_extended_MLE()` checks both conditions for a specified hierarchical model. It returns 0 if both conditions are satisfied, 1 for failure of the Fienberg--Rinaldo existence condition, 2 for failure of identifiability, and 3 if both fail.


The model containing all three two-list interactions is nonidentifiable:

```{r}
MultipleSystemsEstimation::check_extended_MLE(
  Artificial_3,
  "[12,13,23]"
)
```

This returns 2.

The model containing only the AB interaction has no extended maximum likelihood estimate:

```{r}
MultipleSystemsEstimation::check_extended_MLE( Artificial_3, "[12,3]")
```

This returns 1.

Finally, the model containing AC and BC, but not AB, passes both tests:

```{r}
MultipleSystemsEstimation::check_extended_MLE( Artificial_3, "[13,23,1,2]")
```

This returns 0.

These examples reproduce the three distinct behaviours displayed in Table 3 of the paper.

# Checking all possible models

Section 2.7 develops an efficient procedure for determining whether the extended maximum likelihood estimate exists for every possible model containing main effects and an arbitrary subset of the two-list interactions.

The key observation is that, once existence has been established for a model, removing an interaction corresponding to an *overlapping* pair cannot destroy existence. Consequently, if there are \(M\) nonoverlapping pairs, it is sufficient initially to investigate only \(2^M\) top-level models containing all the overlapping pairs.

The original `checkallmodels()` routine implemented this search. It is part of the historical computational machinery for Chan, Silverman and Vincent (2021) and is retained in the now-deprecated `SparseMSE` package rather than in `MultipleSystemsEstimation`.

For the full New Orleans data there are eight lists and 28 possible two-list effects, of which 18 correspond to nonoverlapping pairs. Thus there are \(2^{28}\) possible two-list models in total, but the initial existence search requires only \(2^{18}\) linear-programming checks.

The paper reports that neither nonexistence nor nonidentifiability arises for any model for these data. The same conclusion is reported for the five-list Western data.

For current analyses there is no need to carry out this global search. The estimation routines in `MultipleSystemsEstimation` check the particular candidate models required by their algorithms using `check_extended_MLE()`.

# The stepwise model-selection procedure

Section 3.2 proposes a forward stepwise procedure.

Starting with the main-effects model, every two-list effect not already included is considered. A candidate is discarded if adding it would lead to nonexistence or nonidentifiability. Among the remaining candidates, the effect with the smallest p-value is selected if that p-value is no greater than a specified threshold.

The paper ultimately recommends the threshold

\[
p=0.02.
\]

The current high-level implementation is `estimate_population_stepwise()`.

For a point estimate without bootstrapping:

```{r}
fit <- estimate_population_stepwise(
  Korea,
  pthresh = 0.02
)

fit$popest
fit$MSEfit
```

The same method is also available from the overall package interface:

```{r}
fit <- estimate_population(
  Korea,
  method = "stepwise",
  pthresh = 0.02
)
```

# Bootstrap inference

Section 3.3 points out that confidence intervals conditional on the model selected from the observed data do not allow for uncertainty due to model selection.

The paper therefore repeats the complete stepwise selection procedure for each bootstrap sample.

The bootstrap samples are obtained by multinomial resampling of the observed capture histories. The BCa acceleration is calculated using a delete-one jackknife, exploiting the fact that only distinct observed capture histories need to be considered explicitly.

The current implementation is:

```{r eval=FALSE}
fit_boot <- estimate_population_stepwise(
  Korea,
  nboot = 1000,
  pthresh = 0.02,
  iseed = 1234
)

fit_boot$popest
fit_boot$BCaquantiles
```

A large number of bootstrap replications should be used for substantive inference. Smaller values are useful when checking code.

# New Orleans: Section 4.1

The New Orleans dataset in Section 4.1 contains eight lists.

```{r}
data(NewOrl)

ncol(NewOrl) - 1
sum(NewOrl[, ncol(NewOrl)])
```

There are 28 possible pairs of lists, of which 18 are nonoverlapping.

Applying the stepwise procedure with \(p=0.02\):

```{r}
NewOrl_fit <- estimate_population_stepwise(
  NewOrl,
  pthresh = 0.02
)

NewOrl_fit$popest
NewOrl_fit$MSEfit
```

The paper reports that the selected model contains the two-list effect DE and gives a total population estimate of **1184**.

The published BCa 95% confidence interval, based on 1000 bootstrap replications, is **(717, 1657)**.

The corresponding calculation with the current package is:

```{r eval=FALSE}
NewOrl_boot <- estimate_population_stepwise(
  NewOrl,
  pthresh = 0.02,
  nboot = 1000
)

NewOrl_boot$BCaquantiles
```

The paper also considers the main-effects-only model, obtained when the threshold is 0.01 or smaller. It reports a point estimate of **997** and a 95% confidence interval of **(644, 1618)**.

This can be examined by setting the threshold to zero:

```{r eval=FALSE}
NewOrl_main <- estimate_population_stepwise(
  NewOrl,
  pthresh = 0,
  nboot = 1000
)

NewOrl_main$popest
NewOrl_main$BCaquantiles
```

# Five-list New Orleans data

Because several individual lists in the New Orleans data are very small, the paper also combines the four smallest lists to obtain a five-list version. This is supplied as `NewOrl_5`.

```{r}
data(NewOrl_5)

NewOrl5_fit <- estimate_population_stepwise(
  NewOrl_5,
  pthresh = 0.02
)

NewOrl5_fit$popest
NewOrl5_fit$MSEfit
```

The paper reports that no two-list parameter is significant even at the 5% level. The point estimate is **1034**, with BCa 95% confidence interval **(589, 1703)**.

The bootstrap calculation is:

```{r eval=FALSE}
NewOrl5_boot <- estimate_population_stepwise(
  NewOrl_5,
  pthresh = 0.02,
  nboot = 1000
)

NewOrl5_boot$BCaquantiles
```

# Western site data: Section 4.2

The Western site data contain five lists.

```{r}
data(Western)

Western_fit <- estimate_population_stepwise(
  Western,
  pthresh = 0.02
)

Western_fit$popest
Western_fit$MSEfit
```

The paper reports that the selected model contains the two-list effect AE and that the resulting point estimate is **2483**.

The reported BCa 95% confidence interval is **(1293, 3670)**.

It can be recalculated using:

```{r eval=FALSE}
Western_boot <- estimate_population_stepwise(
  Western,
  pthresh = 0.02,
  nboot = 1000
)

Western_boot$BCaquantiles
```

# Comparison with the BIC approach: Section 3.4

Section 3.4 compares the stepwise/BCa procedure with the BIC approach then implemented in `Rcapture`.

The simulation starts from the model selected by the stepwise procedure for the five-list UK data. The fitted population size and capture-history probabilities are treated as the parameters of a simulated population, and 500 datasets are generated.

For every simulated dataset:

- the BIC procedure selects a model and calculates the confidence interval available in `Rcapture`;
- the stepwise procedure selects a model afresh;
- a BCa interval is constructed for the stepwise estimate, repeating the selection within its bootstrap samples.

The original computational routines used for this simulation are retained in the now-deprecated `SparseMSE` package rather than in `MultipleSystemsEstimation`. `SparseMSE` also retains other routines used for full reproduction of the computational and supplementary material associated with the 2021 paper.

The paper reports population-scale root mean square errors of **3057** for the stepwise method and **5834** for BIC. On the logarithmic scale the corresponding values are **0.19** and **0.34**.

The paper also reports that the nominal 95% BCa interval covered the true value in **90%** of the 500 simulations, compared with **61.4%** for the `Rcapture` method. For nominal 80% intervals, the corresponding coverage rates were approximately **70%** and **42.8%**.

The purpose of this simulation was not to establish universal superiority of one procedure, but to demonstrate the importance of allowing model-selection uncertainty to enter the inference.

# Choosing the threshold: Section 4.3

The paper uses a larger simulation study to investigate the choice of the stepwise threshold.

Seven datasets are used:

- the UK data, in full and five-list versions;
- the Netherlands data, in full and five-list versions;
- the New Orleans data, in full and five-list versions;
- the Western site data.

For each dataset four generating models are considered:

- the full two-list model;
- the main-effects-only model;
- the model selected at threshold 0.001;
- the model selected at threshold 0.05.

This gives 28 simulation scenarios.

For each scenario the paper generates 1000 datasets and obtains estimates using thresholds

\[
0,\ 0.001,\ 0.002,\ 0.005,\ 0.01,\ 0.02,\ 0.05,\ 0.1,\ 1,
\]

where zero corresponds to the main-effects model and one to the full two-list model.

Accuracy is assessed using the mean squared error of the logarithm of the estimated dark figure. Because the overall level of mean squared error differs substantially between scenarios, the paper combines the results by averaging the logarithm of the mean squared error across the 28 scenarios.

The minimum overall score occurs at

\[
\boxed{p=0.02},
\]

which motivates the default threshold used by the package.

The full threshold simulation is deliberately not run as part of the vignette. It involves 28 generating scenarios, 1000 datasets per scenario, and repeated model fitting across multiple thresholds, so it is better regarded as a published simulation study than as an ordinary package example.

The original routines used to carry out this simulation study are retained in the now-deprecated `SparseMSE` package for full reproducibility of the paper.

# Relation to the current package

The original `SparseMSE` package accompanied Chan, Silverman and Vincent (2021). It is now deprecated for current analysis, but is retained as the historical reproducibility package for the original computational and supplementary material. The present `MultipleSystemsEstimation` package retains the methodology while providing a more unified interface.

For a new analysis using the 2021 method, the recommended call is now

```{r}
estimate_population(
  Korea,
  method = "stepwise"
)
```

or, equivalently,

```{r}
estimate_population_stepwise(Korea)
```

The default package interface is `estimate_population(..., method = "auto")`; this uses the BIC-based approach for up to five lists and the stepwise approach for six or more. The main user guide discusses the current choice of methods in detail.

The purpose of the present vignette is different: it records the connection between the current implementation and the calculations in the 2021 paper.

# References

Chan, L., Silverman, B. W. and Vincent, K. (2021).
*Multiple Systems Estimation for Sparse Capture Data: Inferential Challenges When There Are Nonoverlapping Lists*.
*Journal of the American Statistical Association*, **116**(535), 1297--1306.
doi:10.1080/01621459.2019.1708748.

DiCiccio, T. J. and Efron, B. (1996).
*Bootstrap Confidence Intervals*.
*Statistical Science*, **11**(3), 189--228.

Fienberg, S. E. and Rinaldo, A. (2012).
*Maximum Likelihood Estimation in Log-Linear Models*.
*The Annals of Statistics*, **40**, 996--1023.
